Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks

### 2015-12-01 manuscript/manuscript_wtc3.Rmd
lma in discussion
new par figure

regression on gases appendix fig

# Summary
* Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves. A lack of empirical data relating photosynthesis to leaf physiolgoical behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (A), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
* Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~. When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves at the same PPFD. 
* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain. We argue that the rapid response of g~m~ with light enables shade leaves to respond quickly to sunflecks. Evidence that g~m~ not only varies within a canopy but can be up-regulated over short time intervals possibly represents a new mechanism underpinning leaf gas exchange responses to light.

@@

\
Rates of photosynthesis in C~3~ plants are limited by the [CO~2~] available for fixation by Rubisco within the chlorplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmoshpehre to the site of carboxylation [@warren2008stand]. This drawdown consists of multiple resistance pathways to CO~2~ diffusion which include CO~2~ diffusion from the atmosphere through stomata (stomatal conductance, g~s~) and then from these intercellular air spaces into the chloroplast (mesophyll conductance, g~m~). Based on optimality theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize A [@peltoniemi2012co]. This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal]. Mesophyll conductance can also impose limitations on A as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in A [@niinemets2007photosynthesis] if g~m~ constrains CO~2~ supply to the chloroplast. Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified. Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack empirical data currently hinders are ability to interpreting their coupled responses to A across sun and shade leaves.

@@

\
3. As shade leaves were expected to develop lower biochemical photosynthetic capacity and leaf physiolgoical responses to light are typically slow (minutes), increases in A following sunfleck simulations were not expected to reach rates of full sun leaves.

\
4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporative demand from higher temperatures will lead to stomatal closure.

@@

## Leaf carbon isotope discrimination and mesophyll conductance
The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at liw light (Figure `r deltafig`). Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments. The remainder $\Delta_o$ consists of the contributions of g~s~, respiration and photorespiration to discrimination.

@@

# Discussion
Here we show that A in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased. Although shade leaves in lower light environments exhibited relatively high g~s~ and c~i~, it was rapid increases in g~m~ under periods of high light availability that led to this upregulation of A. Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain. These findings offer new insights into how aspects of leaf physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake in future studies. Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology.

## Resource distribution and photosynthetic capacity
The allocation of N~a~ constrains A and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain. Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979). As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical]. Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and A were all reduced in shade leaves.

@@

**Figure 1**. Local light environment for sun and shade leaves for each monthly gas exchange campaign. Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf.

@@

\
**Figure 6**. Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at hight light and shade leaves measured at both low and high light. The solid line represents the theoretical line for C3 plants from Evans et al. [-@evans1986carbon].

@@

\
**Figure S3**. Response of A (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign.

\
**Figure S4**. Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign.

@@

# Figures {.unnumbered}
```{r par, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/par.R")
```

@@

```{r gasex_temp1, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/gasex_temp.R")
```

@@

```{r gasex_temp2, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/gasex_temp2.R")
```

>>>>

# Summary
* Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves. A lack of empirical data relating photosynthesis to leaf physiological behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (A), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
* Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~. When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves at the same PPFD. 
* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain. We argue that the rapid response of g~m~ with light enables shade leaves to respond quickly to sunflecks. Evidence that g~m~ not only varies within a canopy but can be up-regulated over short time intervals possibly represents a new mechanism underpinning leaf gas exchange responses to light.

@@

\
Rates of photosynthesis in C~3~ plants are limited by the [CO~2~] available for fixation by Rubisco within the chloroplast and this [CO~2~] is a function of the drawdown of CO~2~ from the atmosphere to the site of carboxylation [@warren2008stand]. This drawdown consists of multiple resistance pathways to CO~2~ diffusion which include CO~2~ diffusion from the atmosphere through stomata (stomatal conductance, g~s~) and then from these intercellular air spaces into the chloroplast (mesophyll conductance, g~m~). Based on optimal theory, regulation of g~s~ within a tree canopy should act to efficiently utilize available supplies of light, N and water to maximize A [@peltoniemi2012co]. This is because stomata are hypothesized to exhibit an optimal behaviour to maximize C gain while simultaneously minimizing water loss through transpiration [@cowan1977stomatal]. Mesophyll conductance can also impose limitations on A as large g~s~ [@warren2008stand; @ubierna2011estimation], reducing the efficiency of leaf N use in A [@niinemets2007photosynthesis] if g~m~ constrains CO~2~ supply to the chloroplast. Part of the variation in photosynthetic capacity between sun and shade leaves has been proposed to be due to differences in g~m~ [@piel2002effect; @warren2007internal], yet the trade-offs that constrain this diffusion pathway are yet to be explicitly quantified. Stomatal and mesophyll conductance should not be considered independent of each other [@griffiths2013mesophyll], but a lack empirical data currently hinders are ability to interpreting their coupled responses to A across sun and shade leaves.

@@

\
3. As shade leaves were expected to develop lower biochemical photosynthetic capacity and leaf physiological responses to light are typically slow (minutes), increases in A following sunfleck simulations were not expected to reach rates of full sun leaves.

\
4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporation demand from higher temperatures will lead to stomatal closure.

@@

## Leaf carbon isotope discrimination and mesophyll conductance
The observed carbon isotope discrimination ($\Delta_o$) measured during photosynthesis was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001), with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light than shade leaves at low light (Figure `r deltafig`). Carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta_o$ (69.7&pm;0.4%) and varied little across measurement temperatures, leaf types, or warming treatments. The remainder $\Delta_o$ consists of the contributions of g~s~, respiration and photorespiration to discrimination.

@@

# Discussion
Here we show that A in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased. Although shade leaves in lower light environments exhibited relatively high g~s~ and c~i~, it was rapid increases in g~m~ under periods of high light availability that led to this up-regulation of A. Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain. These findings offer new insights into how aspects of leaf physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake in future studies. Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology.

## Resource distribution and photosynthetic capacity
The allocation of N~a~ constrains A and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain. Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979). As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical]. Our data agree with these conventional conclusions as the distribution of N~a~, both measures of photosynthetic capacity and A were all reduced in shade leaves. Leaf mass area, however, was not different between sun and shade leaves. This could be due to leaf formation under comparative light conditions or possible differences in total non-structural carbohydrates contents between leaf types.

@@

**Figure 1**. Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014. Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf. Each date represents the starting date for each measurement campaign. Points represent the maximum daily air temperature during each campaign period.

@@

\
**Figure 6**. Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light. The solid line represents the theoretical line for C3 plants from Evans et al. [-@evans1986carbon].

@@

\
**Figure S3**. Response of A (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign. Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns. All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination. Weak negative relationships with g~m~ and increasing leaf temperature were detected with sun and shade leaves under their local light environment (R^2^ = 0.16 and 0.08, respectively).

\
**Figure S4**. Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light. Shaded symbols represents each monthly measurement campaign. Solid lines, colored by leaf and light type, are fitted line for the relationship with each parameter and leaf temperature across all measurement campaigns. All parameters with no relationship are fitted with zero slope and the overall mean value for each treatment combination. Leaf VPD inside the gas exchange cuvette was positively correlated with increasing leaf temperature for sun leaves and shade leaves at low and high light (R^2^ = 0.73, 0.58 and 0.72, respectively)

@@

# Figures {.unnumbered}
```{r par, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/par2.R")
```

@@

```{r gasex_temp3, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/gasex_temp3.R")
```

@@

```{r gasex_temp4, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=400, dev.args=list(type="cairo")}
source("master_scripts/gasex_temp4.R")
```
